### 2016-03-10 docs/ms_communitySynchrony.Rmd
restructures ms to close #20

\textbf{Number of Tables and Figures}:  2 Tables;  4 Figures

@@

<!--
One potential way forward is to use population models fit to empirical data to understand the influence of different drivers on species synchrony.
@Mutshinda2009 used such an approach by fitting a dynamic population model to several community time series of insect and bird abundances.
They used a statistical technique to decompose temporal variation into competition and environmental components, and found that postively correlated environmental responses among species determined community dynamics.
While a major step forward, Mutshinda et al.'s (2009) modeling technique relied on abundance data that may or may not reliably capture competitive interactions that occur at the individual level.
Despite almost a decade of research aimed toward understanding species synchrony, we still lack direct tests that link environmental fluctuations, interspecific interactions, and demographic stochasticity to species synchrony in natural communities.

Here, we directly test the importance of proposed drivers of species synchrony by parameterizing and simulating dynamic, multi-species population models fit to long term demographic data from five semi-arid grasslands [@Chu2015].
We use the fitted population models to simulate communities with and without the three mechanisms identified by theory as drivers of species synchrony: interspecific interactions, environmental forcing, and demographic stochasticity [@Loreau2008a; @Loreau2013; @DeMazancourt2013a].
Our modeling approach allows us to draw inference from real ecological systems that would otherwise be impossible using strictly statistical [e.g., @Mutshinda2009], experimental [e.g., @Gross2014], or observational [e.g., @Houlahan2007] approaches.
Given that our models capture the essential features of community dynamics important to synchrony (density-dependence, demographic and environmental stochasticity) and that these models are able to reproduce observed plant abundances [@Chu2015], perturbing the models reveals the processes responsible for the degree of species synchrony in our focal grassland communities.
Our main focus is on the synchrony of species' per capita growth rates since they represent a species immediate response to environmental conditions and competition.
However, we also analyze synchrony of plant cover (a metric of abundance) to relate our findings to previous research.
-->

@@

Using the above methods, we ran six simulation experiments:
(1) All drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) present using the IBM;
(2) Demographic stochasticity removed using the IPM;
(3) Environmental stochastcity removed suing the IBM;
(4) Interspecific competition removed using the IBM;
(5) Interspecific competition and demographic stochasticity removed using the IPM;
(6) Interspecific competition and environmental stochasticity removed using the IBM.
We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
This allowed species time to reach their stable size distribution.
We then calculated the synchrony of species' per capita growth rates over 100 randomly selected contiguous 50 time-step sections.
We ran IBM simulations for 100 time steps, and repeated the simulations 100 times for each simulation experiment.
From those, we retained only the simulations in which no species went extinct due to demographic stochasticity.
Synchrony was calculated over the 100 time steps for each no extinction run within a model experiment.
To explore the effect of demographic stochasticity in different sized populations, we ran simulations (1) and (6) on plot sizes of 1, 2, 3, 4, and 5 square meters.

@@

DISCUSSION
----------
Our study produced three main findings that were generally consistent across five natural plant communities: (1) interspecific interactions were weak and therefore had little impact on community synchrony; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; and (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated.
These results are in agreement with our theoretical prediction for species synchrony under the limiting case of community dynamics being driven by environmental fluctuations alone.

@@

### Demographic stochasticity decreases synchrony
Our theoretical prediction for species synchrony in a neutral community with no species interactions or environmental fluctuations ($\mathcal{M}_D$) received no support when parameterized with emprical estimates (Fig. 1).
Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small (Fig. 3), in agreement with theory [@Loreau2008a].
Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations (Fig. 3).
These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
Only in the Idaho community does synchrony of per capita growth rates approach independent fluctuations (1/*S*, where *S* is the number of specuies) in a non-constant environment (Fig. 3).
This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.

### Environmental fluctuations drive community synchrony
In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations (Equation 4).
Under such conditions community synchrony should approximately equal the synchrony of species' responses to the environment [@Loreau2008a].
Several lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
First, of our two limiting-case theoretical predictions for species synchrony (Equations 3-4), only our prediction for communities driven by environmental fluctuations alone ($\mathcal{M}_E$) received empirical support (Fig. 1).
Second, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
Lastly, removing environmental fluctuations from simulations consistently reduced synchrony (Fig. 2), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig. 3).
Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony.
Likewise, our results appear to be quite general: across four arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
Only in the Idaho community did not removing environmental fluctuations not cause a large decrease in synchrony, but removing competition had no effect either.
Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural plant communities.

Another challenge to the generality of our results is a recent analysis of several biodiversity-ecosystem function experiments that came to a completely opposing conclusion: competition drives species synchrony in grasslands [@Gross2014].
@Gross2014 found that temporal asynchrony among plant species in four experiments was driven by interspecific competition that counteracted the synchronizing effect of environmental fluctuations.
This is surprising given the extremely weak species interactions we estimated.
But the apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
Gross et al.'s results are based on experimentally-assembled communities that have coexisted for, at most, 11 years.
In contrast, the natural communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap [@Chu2015].
In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments community assembly processes that can result in competitive exclusion may yet to have been realized.
Therefore, such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
This also leads to the prediction that community synchrony is driven by competition in recently assembled communities, while synchrony is driven by environmental fluctuations in communities that have undergone extensive species sorting.
In other words, in the absence of disturbance and invasion, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.

<!--
### The importance of functional diversity
Our analysis also contributes to mounting evidence on the importance of functional diversity for maintaining temporal stability via species asynchrony [@Reich2012; @Polley2013].
Functional traits define how a species will respond to its environment.
Therefore, greater functional differences among species should lead to lower correlations in species' responses to the environment.
We found evidence for the negative effect of functional diversity on community synchrony: community synchrony was lowest Idaho (Table 2, Fig. 2); the only community that contained two life forms (one shrub, three perennial grasses).
In fact, when we ran simulations with the shrub removed, synchrony of the Idaho community increased substantially and was similar to synchrony in the Kansas community (Fig. SX), which also contained three species.
Functional diversity is especially important in communities with weak interspecific interactions where synchrony depends soley on species' environmental responses.
Adding species to such communities will always decrease synchrony, and increase temporal stability, so long as the synchrony of species' environmental response is less than 1 [@Loreau2008a; @Loreau2013].
But functionally diverse suites of species will have the lowest synchrony of environmental responses, and thus lowest community synchrony and highest stability.
-->

@@

table_data <- as.data.frame(rbind(get_table_metrics(out_ks),
                                  get_table_metrics(out_id),
                                  get_table_metrics(out_mt),
                                  get_table_metrics(out_nm),
                                  get_table_metrics(out_az)))
table_data$Site <- c("Kansas", "Idaho", "Montana", "New Mexico", "Arizona")
table_data <- table_data[,c("Site", "V1", "V4", "V7")]
table_data$spp <- c(length(unique(ks_data$species)),
                    length(unique(id_data$species)),
                    length(unique(mt_data$species)),
                    length(unique(nm_data$species)),
                    length(unique(az_data$species)))
colnames(table_data) <- c("Site",
                          "$\\phi_{r}$", 
                          "$\\phi_{C}$",
                          "CV of Total Cover",
                          "Species richness")
# colnames(table_data) <- c("Site",
#                           "$\\phi_{r}$", 
#                           "$\\text{E}(\\phi_{r})$", 
#                           "$\\text{E}(\\sigma_{r,c} + \\sigma_{r,e})$",
#                           "$\\phi_{C}$", 
#                           "$\\text{E}(\\phi_{C})$",
#                           "$\\text{E}(\\sigma_{C,c} + \\sigma_{C,e})$",
#                           "CV of Total Cover",
#                           "Species richness")
# table_data[,3] <- round(1/table_data[,ncol(table_data)],2) # reset expected sybchrony to 1/S
# table_data[,4] <- table_data[,2] - table_data[,3] # calculate obs - exp synchrony

@@

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/formatted_fig1.png}
  \caption{Observed versus predicted synchrony of growth rates under our two limiting case predictions. The observed synchrony of per capita growth rates was best predicted by our theoretical prediction in the limiting case of only environmental stochasticity operating ($\phi_R$; green line and points in left panel): it was the only regression line whose slope and intercept did not significantly differ from zero or one, respectively.}
\end{figure}

@@

<!--
\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/ibm_sims_fig.png}
  \caption{Community synchrony of per capita growth rates (top panels) and percent cover (bottom panels) from spatially-explicit IBM simulations. Points represent replicate IBM simulations (out of 100 possible) in which no species went extinct. Lines show linear fits between plot size (proxy for population size) and species synchrony. Demographic stochasticity plays a large role in small populations (small plot sizes) and a smaller role in large populations (large plot sizes). Simulation type codes: C = constant environment, F = fluctuating environment, INTER = interspecific interactions included, NoINTER = interspecific interactions excluded (set to zero).}
\end{figure}
-->

<!--
\pagebreak{}

\begin{figure}[!ht]
  \centering
      \includegraphics{./components/envresp_vs_ipmsynch.png}
  \caption{Average synchrony of IPM-simulated per capita growth rates (from 13 year sections) as a function of the average synchrony of the vital rate random year effects for each site. Community sychrony is expected to approximately equal the synchrony of species' environmental responses in the absence of demographic stochasticity and strong interspecific interactions. The synchrony of species' random year effects across vital rates is a strong ($\text{R}^2$ = 0.67), but not perfect, predictor of community synchrony. Points are labeled by site codes: AZ = Arizona, ID = Idaho, KS = Kansas, MT = Montana, and NM = New Mexico.}
\end{figure}
-->
\pagebreak{}

>>>>

\textbf{Number of Tables and Figures}:  3 Tables;  3 Figures

@@

@@

Using the above methods, we ran six simulation experiments:
(1) All drivers (environmental stochasticity, demographic stochasticity, or interspecific competition) present using the IBM;
(2) Demographic stochasticity removed using the IPM;
(3) Environmental stochastcity removed suing the IBM;
(4) Interspecific competition removed using the IBM;
(5) Interspecific competition and demographic stochasticity removed using the IPM; and
(6) Interspecific competition and environmental stochasticity removed using the IBM.
We ran IPM simulations for 2,000 time steps, after an initial 500 iteration burn-in period.
This allowed species time to reach their stable size distribution.
We then calculated the synchrony of species' per capita growth rates over 100 randomly selected contiguous 50 time-step sections.
We ran IBM simulations for 100 time steps, and repeated the simulations 100 times for each simulation experiment.
From those, we retained only the simulations in which no species went extinct due to demographic stochasticity.
Synchrony was calculated over the 100 time steps for each no extinction run within a model experiment.
To explore the effect of demographic stochasticity in different sized populations, we ran simulations (1) and (6) on plot sizes of 1, 2, 3, 4, and 5 square meters.

@@

DISCUSSION
----------
Our study produced four main findings that were generally consistent across five natural plant communities: (1) limiting-case predictions from the theoretical model were well-supported by analagous simulations from the empirical models; (2) demographic stochasticity decreased community synchrony, as expected by theory, and its effect was largest in small populations; (3) environmental fluctuations increased community synchrony relative to simulations in constant environments because species-specific responses to the environment were positively, but not completely, correlated; and (4) interspecific interactions were weak and therefore had little impact on community synchrony.

### Simulations support theoretical predictions

### Demographic stochasticity decreases synchrony
Demographic stochasticity is able to partially counteract the synchronizing effects of environmental fluctuations and interspecific interactions on per capita growth rates, but only when populations are small (Fig. 3), in agreement with theory [@Loreau2008a].
Even in small populations, however, demographic stochasticity is not a strong enough force to induce independent fluctuations (Fig. 3).
These results confirm the theoretical argument by @Loreau2008a that independent fluctuations among interacting species in a non-constant environment should be rare.
Only in the Idaho community does synchrony of per capita growth rates approach independent fluctuations (1/*S*, where *S* is the number of specuies) in a non-constant environment (Fig. 3).
This is most likely due to the strong effect of demographic stochasticity on the shrub *Artemisia tripartita* since even a 5 $\text{m}^2$ quadrat would only contain a few individuals of that species.

### Environmental fluctuations drive community synchrony
In large populations where interspecific interactions are weak, synchrony is expected to be driven exclusively by environmental fluctuations (Equation 4).
Under such conditions community synchrony should approximately equal the synchrony of species' responses to the environment [@Loreau2008a].
Two lines of evidence lead us to conclude that environmental fluctuations drive species synchrony in our focal plant communities.
First, in our model simulations, removing interspecific interactions resulted in no discernable change in community-wide species synchrony of per capita growth rates (Fig. 2).
Second, removing environmental fluctuations from simulations consistently reduced synchrony (Fig. 2), and the synchrony of species in monoculture is a very strong predictor of synchrony of species in polyculture (Fig. 3).
Our results lead us to conclude that environmental fluctuations, not species interactions, are the primary driver of community-wide species synchrony.
Likewise, our results appear to be quite general: across four arid-to-subhumid grasslands community synchrony is almost soley due to environmental fluctuations since interspecific interactions are weak.
Only in the Idaho community did removing environmental fluctuations not cause a large decrease in synchrony, but removing competition had no effect either.
Given accumulating evidence that niche differences in natural communities are large [reviewed in @Chu2015], and thus species interactions are likely to be weak, our results may be general in natural plant communities.

@@

Our results are in direct contrast to a recent analysis of several biodiversity-ecosystem function experiments that found competition drives species synchrony in grasslands [@Gross2014].
This is surprising given the extremely weak species interactions we estimated.
But the apparent inconsistency between our results and those of @Gross2014 disappears when we consider the differences between our studies.
Gross et al.'s results are based on experimentally-assembled communities that have coexisted for, at most, 11 years.
In contrast, the natural communities we studied consist of species that have coexisted for decades and have virtually zero niche overlap [@Chu2015].
In theory, species interactions should weaken through time as community assembly progesses [@Kokkoris1999], meaning that in biodiversity-ecosystem function experiments community assembly processes that can result in competitive exclusion may yet to have been realized.
Therefore, such communities may be highly influenced by transient, but strong, interspecific competition that can mask the synchronizing effects of the environment.
This also leads to the prediction that community synchrony is driven by competition in recently assembled communities, while synchrony is driven by environmental fluctuations in communities that have undergone extensive species sorting.
In other words, in the absence of disturbance and invasion, the dominant driver of community synchrony should shift from competition to environmental fluctuations through time.
One way to test this prediction is to continue collecting data from biodiversity-ecoystem function experiments and conduct the analysis of @Gross2014 every few years.

@@

table_data <- as.data.frame(rbind(get_table_metrics(out_az),
                                  get_table_metrics(out_id),
                                  get_table_metrics(out_ks),
                                  get_table_metrics(out_mt),
                                  get_table_metrics(out_nm)))
alltable_data <- table_data
table_data$Site <- c("Arizona", "Idaho", "Kansas", "Montana", "New Mexico")
table_data <- table_data[,c("Site", "V1", "V4", "V7")]
table_data$spp <- c(length(unique(az_data$species)),
                    length(unique(id_data$species)),
                    length(unique(ks_data$species)),
                    length(unique(mt_data$species)),
                    length(unique(nm_data$species)))
colnames(table_data) <- c("Site",
                          "$\\phi_{r}$", 
                          "$\\phi_{C}$",
                          "CV of Total Cover",
                          "Species richness")
# colnames(table_data) <- c("Site",
#                           "$\\phi_{r}$", 
#                           "$\\text{E}(\\phi_{r})$", 
#                           "$\\text{E}(\\sigma_{r,c} + \\sigma_{r,e})$",
#                           "$\\phi_{C}$", 
#                           "$\\text{E}(\\phi_{C})$",
#                           "$\\text{E}(\\sigma_{C,c} + \\sigma_{C,e})$",
#                           "CV of Total Cover",
#                           "Species richness")
# table_data[,3] <- round(1/table_data[,ncol(table_data)],2) # reset expected sybchrony to 1/S
# table_data[,4] <- table_data[,2] - table_data[,3] # calculate obs - exp synchrony

@@

```{r prediction_table, echo=FALSE, include=TRUE, results='asis', message=FALSE, eval=T, cache=FALSE}
exp_synchrony_env <- readRDS("../results/envstoch_predictions.RDS")
exp_synchrony_env <- exp_synchrony_env[complete.cases(exp_synchrony_env),]
exp_synchrony_demo <- alltable_data[,c("V2","V5")]
exp_synchrony_demo$site <- table_data$Site
colnames(exp_synchrony_demo) <- c("pgr", "cover", "site")
sims_output <- readRDS("../results/allsims_plot_data.RDS")

pred_table <- data.frame(site = table_data$Site,
                         exp_demo = exp_synchrony_demo$pgr,
                         sim_demo = round(subset(sims_output, simulation=="6No Comp. + No E.S.")["synchrony"],2),
                         exp_env = round(exp_synchrony_env$growthrate_prediction,2),
                         sim_env = round(subset(sims_output, simulation=="5No Comp. + No D.S.")["synchrony"],2))

colnames(pred_table) <- c("Site", 
                          "Predicted $\\phi_{R,\\mathcal{M}_D}$",
                          "Simulated $\\phi_{R,\\mathcal{M}_D}$",
                          "Predicted $\\phi_{R,\\mathcal{M}_E}$",
                          "Simulated $\\phi_{R,\\mathcal{M}_E}$")

pred_cap <- "Predicted and simulated synchrony of per capita growth rates under the two limiting cases."
print(xtable(pred_table, caption = pred_cap),
      caption.placement="top",
      include.rownames = F, 
      sanitize.colnames.function = identity,
      comment=FALSE,
      size="normalsize")
```

@@

<!--
\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/formatted_fig1.png}
  \caption{Observed versus predicted synchrony of growth rates under our two limiting case predictions. The observed synchrony of per capita growth rates was best predicted by our theoretical prediction in the limiting case of only environmental stochasticity operating ($\phi_R$; green line and points in left panel): it was the only regression line whose slope and intercept did not significantly differ from zero or one, respectively.}
\end{figure}

\pagebreak{}
-->

@@
